The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-10-05 article.tex
More tweaks (it's actually a pretty good read)

The open source cheminformatics community has made significant steps
forward recently~\cite{OBoyle2011b} as evidenced by the growing number
of tools and underlying toolkits, along with the usage of these
software components in a variety of applications.
The Chemistry Development Kit (CDK) is one of the tools developed
under the aegis of the Blue Obelisk 
and previously documented versions have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
Use of the CDK ranges from inclusion of CDK functionality in
wrapper platforms such as Cinfony~\cite{OBoyle2008}, incorporation
within the R environment (rcdk~\cite{Guha2007}), and as plugins for
Taverna~\cite{Truszkowski2011}, 
KNIME~\cite{Beisken2013}, Cytoscape (ChemViz2~\cite{ChemViz2}), and for
Microsoft Excel (LICSS~\cite{Lawson2012}).
In contrast to scenarios that have made CDK functionality available in
larger systems, a number of projects have employed the CDK as a
general cheminformatics toolkit. Examples include 
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
PaDEL~\cite{yap2011padel}, ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014,Rahman2016},
WhichCyp~\cite{Rostkowski2013}, MetaPrint2D~\cite{Carlsson2010}, and the IUPHAR/BPS Guide to Pharmacology database~\cite{Southan2016}.
A number of such tools were initially developed using older versions
of the CDK and are updated to new releases as they are made
available. Examples include Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and
AMBIT~\cite{jeliazkova2011ambit,jeliazkova2011ambitsmarts,kochev2013ambit}. The
CDK has also played a role in a number of chemical studies, such as finding
the maximally bridging rings in chemical structures~\cite{Marth2015}, and
prediction of organic reactions~\cite{Segler2016}.

@@

  In previous versions of the API, stereochemistry was represented disparately hindering 
  interconversion between and within file formats. In CDK \cdkversion{} the core representation 
  has been standardized upon and procedures updated or added to enable duplicate checking,
  pattern matching, and interconversion.

@@

\vspace{0.2cm}
  \begin{itemize}
   \item perception from 2D coordinates;
   \item perception from 3D coordinates;
   \item wedge assignment;
   \item graph (sub)isomorphism matching;
   \item SMARTS matching; and
   \item canonicalization.
  \end{itemize}
\vspace{0.2cm}

@@

  \subsubsection*{Signatures}

@@

  Finally, a signature fingerprint can be calculated for molecules, to allow
  similarity calculations. This can then be used in QSAR modeling~\cite{signaturefingerprints,
  Spjuth2011DS,Moghadam2015,Alvarsson2014,Spjuth2012OS,Norinder2013,Alvarsson2016}.

@@

  Many of the rendering options are available as parameters in the
  core API and as methods on the DepictionGenerator class. This
  includes substructure coloring, exemplified with an example reaction
  shown in Figure~\ref{fig:depiction}.

>>>>

The open source cheminformatics community has made significant steps
forward recently~\cite{OBoyle2011b} as evidenced by the growing number
of tools and underlying toolkits, along with the usage of these
software components in a variety of applications.
The Chemistry Development Kit (CDK) is one of the tools developed
under the aegis of the Blue Obelisk 
and previously documented versions have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
Use of the CDK ranges from inclusion of CDK functionality in
wrapper platforms such as Cinfony~\cite{OBoyle2008}, incorporation
within the R environment (rcdk~\cite{Guha2007}), and as plugins for
Taverna~\cite{Truszkowski2011}, 
KNIME~\cite{Beisken2013}, Cytoscape (ChemViz2~\cite{ChemViz2}), and for
Microsoft Excel (LICSS~\cite{Lawson2012}).
In contrast to scenarios that have made CDK functionality available in
larger systems, a number of projects have employed the CDK as a
general cheminformatics toolkit. Examples include 
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
PaDEL~\cite{yap2011padel}, ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014,Rahman2016},
WhichCyp~\cite{Rostkowski2013}, MetaPrint2D~\cite{Carlsson2010}, and the IUPHAR/BPS Guide to Pharmacology database~\cite{Southan2016}.
A number of such tools were initially developed using older versions
of the CDK and are updated to new releases as they are made
available. Examples include Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and
AMBIT~\cite{jeliazkova2011ambit,jeliazkova2011ambitsmarts,kochev2013ambit}. The
CDK has also played a role in a number of chemical studies, such as finding
the maximally bridging rings in chemical structures~\cite{Marth2015},
prediction of organic reactions~\cite{Segler2016}, and bioactivities
of compounds~\cite{Alvarsson2016}.

@@

  Previous versions of the API presented stereochemistry disparately hindering 
  interconversion between and within file formats. CDK \cdkversion{} standardizes
  upon a new core representation and procedures have been updated or added to
  enable duplicate checking, pattern matching, and interconversion.

@@

\vspace{0.2cm}
  \begin{itemize}
   \item perception from 2D coordinates,
   \item perception from 3D coordinates,
   \item wedge assignment,
   \item graph (sub)isomorphism matching,
   \item SMARTS matching, and
   \item canonicalization.
  \end{itemize}
\vspace{0.2cm}

@@

  \subsubsection*{Atomic and Molecular Signatures}

@@

  Finally, a signature fingerprint can be calculated for molecules, to allow
  similarity calculations. This can then be used in QSAR modeling~\cite{Alvarsson2016,
  signaturefingerprints,Spjuth2011DS,Moghadam2015,Alvarsson2014,Spjuth2012OS,Norinder2013}.

@@

  Many of the rendering options are available as parameters in the
  core API and as methods on the DepictionGenerator class. This
  includes substructure coloring, exemplified with an example reaction
  shown in Figure~\ref{fig:depiction}.
  When missing, 2D coordinates are generated on the fly with the new structure
  diagram layout functionality.
